Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Fixed typos.

\subsection{Full hydration: Experimental order parameters for the glycerol backbone and headgroup}\label{experiments}
The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
allowing the absolute value order parameter measurements for these segments by $^2$H NMR~\cite{gally75,brown77,brown78,akutsu81}.
In addition, the absolute values of order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
DMPC~\cite{hong95b,gross97,dvinskikh05a}, DOPC~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
have been measured with several different implementations of $^1$H-$^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
are measured with $^1$H-$^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13} with the signs measured in~\cite{hong95a,hong95b,gross97} are shown in Fig.~\ref{HGorderparameters}.
\begin{figure}[]
%  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
\newline
  \caption{\label{HGorderparameters}
  Order parameteres from simulations listed in Table~\ref{systems} and experiments for glycerol and choline groups.
The experimental values were taken from the following publications:
 DMPC 303~K from \cite{gross97},
 DMPC 314~K from \cite{dvinskikh05a},
 DPPC 322~K from \cite{gally75},
 DPPC 323~K from \cite{akutsu81},
 POPC 296~K from \cite{bechinger91}, and
 POPC 300~K from \cite{ferreira13}.
  The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
these systems we had at least two data sets (see Table~\ref{systems});
the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average. 
The interactive version of this figure is available at  https://plot.ly/$\sim$HubertSantuz/72/lipid-force-field-comparison/.
} 
\end{figure}

In general there is a good agreement between the order parameters measured with different experimental NMR techniques: Almost all the 
reported values are within a variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
for all fully hydrated PC bilayers, regardless of the variation in their acyl chain composition and temperature.
Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
These experiments are not shown in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
or the spectral resolution is quite low and the numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
Due to this end, it is highly likely that these reported lower order parameters are due to lower experimental 
accuracy and therefore we exclude them from our discussion. 
For more details, see Ref.~\citenum{accuracyPOST}. 
Motivated by the high experimental reproducibility, we have highlighted in 
Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas spanning 0.04 units), within which we expect the calculated absolute 
values of the order parameters of a well-performing force field to fall.

>>>>

\subsection{Full hydration: Experimental order parameters for the glycerol backbone and headgroup}\label{experiments}
The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
allowing the absolute value order parameter measurements for these segments by $^2$H NMR~\cite{gally75,brown77,brown78,akutsu81}.
In addition, the absolute values of order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
DMPC~\cite{hong95b,gross97,dvinskikh05a}, DOPC~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
have been measured with several different implementations of $^1$H-$^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
are measured with $^1$H-$^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13} with the signs measured in~\cite{hong95a,hong95b,gross97} are shown in Fig.~\ref{HGorderparameters}.
\begin{figure}[]
%  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
\newline
  \caption{\label{HGorderparameters}
  Order parameteres from simulations listed in Table~\ref{systems} and experiments for glycerol and choline groups.
The experimental values were taken from the following publications:
 DMPC 303~K from \cite{gross97},
 DMPC 314~K from \cite{dvinskikh05a},
 DPPC 322~K from \cite{gally75},
 DPPC 323~K from \cite{akutsu81},
 POPC 296~K from \cite{bechinger91}, and
 POPC 300~K from \cite{ferreira13}.
  The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
these systems we had at least two data sets (see Table~\ref{systems});
the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average. 
An interactive version of this figure is available at  https://plot.ly/$\sim$HubertSantuz/72/lipid-force-field-comparison/.
} 
\end{figure}

In general there is a good agreement between the order parameters measured with different experimental NMR techniques: Almost all the 
reported values are within a variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
for all fully hydrated PC bilayers, regardless of the variation in their acyl chain composition and temperature.
Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
In these experiments, however, the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
or the spectral resolution is quite low and numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
As it, therefore, is highly likely that the reported lower order parameters are due to lower experimental 
accuracy, we exclude them from the present discussion;
for more details, see Ref.~\citenum{accuracyPOST}. 
Motivated by the high experimental reproducibility, we have highlighted in 
Fig.~\ref{HGorderparameters} subjective sweet spots (light blue areas spanning 0.04 units), within which we expect the calculated  
values of the order parameters of a well-performing force field to fall.
